Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-23 TACAS2018/paper.tex
moved quantization noise to appendix

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Errors generated by numerical representations} 
\label{sec:numeric_rep}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The models we consider have two sources of error that are due to numerical  
representation which we commonly refer to as FWL effects.
The first is the numerical error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$, $C_d$, $D_d$, $x_k$ and $y_k$.
The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
representation of numbers, and briefly describe the errors
introduced.  A~formal discussion is in
Appendix~\ref{sec:numeric_rep}.

Let $\mathbb{R}_{\langle M,F \rangle}$ be a fixed point domain with $M$ bits
representing the binary word length of the representation and $F$ bits
representing the prescaling $2^{-F}$ (\emph{i.e} the number of bits after
the decimal place).  We note that a negative $F$ denotes an integer
multiplication that adds $-F$ trailing zeros in the number represented by
$M$.
%Let also $\mathbb{R}_{\langle M,[E] \rangle}$ be a floating point
%domain with $M$ bits mantissa and $E$ bits exponent.  This domain
%corresponds to $$\mathbb{R}_{\langle M,[E] \rangle} = \bigcup_F
%\mathbb{R}_{\langle M,F \rangle} \mid F \in [2^{1-E}, 2^{E-1}]$$ This
%definition important because it means that the distance between any two
%numbers in the domain depends on the location of those numbers.  Since
%a floating point domain is a union of a number of fixed
%point domains, we will continue our reasoning on fixed point only, which
%will apply to the floating point domain by extension.  Operations on these
%numbers will incur errors that can be bound
%by~\cite{DBLP:conf/arith/BrainTRW15}.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Modeling quantization as noise} \label{sec:quantization-noise}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\note{Add a summary sentence and move the rest to the appendix.}
In Figure \ref{fig:observersystem2} we introduced
several sources of noise, which are due to quantization effects. We now describe
their nature and create a sound model for them to be used during our synthesis
and verification process.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Round-off errors}
\label{sec:roundoff-noise}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

%% %-------------------------------
%% \section{Specification of control systems properties} 
%% \label{sec:specification}
%% %-------------------------------

%% %-------------------------------
%% \subsection{Stability} 
%% \label{ssec:stabspecification}
%% %-------------------------------

%% %-------------------------------
%% \subsection{Safety} 
%% \label{ssec:safespecification}
%% %-------------------------------

@@

Since we are interested in synthesising controllers with observer
designs, in this section we will look at checking stability, safety
and the response time in the presence of an observer. A full description of
the theory related to the observer design is provided in appendix \ref{sec:control_theory}.

@@

The ability of a system to remain within a bounded set of states is called stability.  
There are different types of stability, such as Bounded Input/Bounded Output (BIBO)
in which the requirement is that given a bounded input the output remains within
certain limits, or asymptothic stability in which the output is expected to
converge to a certain value. In our model, stability is directly related to
convergence, which means it directly relates to the eigenvalues of the dynamics. 
In this paper we use Jury's criterion~\ref{sec:closed_stability} to verify stability.

@@

The benchmark suite employed to evaluate the proposed technique is 
composed by 7 SISO models extracted from the literature~\cite{acrobot,cstr,CHEN1979389,KOKOTOVIC198023,gajic2008optimal,Franklin15,maglev,converters}.
In order to obtain statistically relevant results, we have taken these models
and modified each possible combination of their parameters by a 10\% tolerance,
thus creating hundreds \note{how many?} of different benchmarks from this base.
In the verification module, we have used the tool Axelerator~\cite{} .

@@

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{count} & \multicolumn{2}{|c|}{CEGIS ($\mat{K}$)} & \multicolumn{5}{|c|}{CEGORS($\mat{K}+\mat{L}+T_s$)} \\
   &                                 & & & \multicolumn{1}{|c|}{-AR1} & \multicolumn{1}{|c|}{-OR} & \multicolumn{1}{|c|}{-AR2} & \multicolumn{1}{|c|}{-OR} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max}\\\hline
1  & DC Motor          & 2 & 10 & 9.5\,s & 2.1\,s& 1.9\,s& 3.4\,s& 1.2\,s &   3.1\,s &   14.1\,s\\
2  & Helicopter        & 3  & 10 & \xmark & 1.4\,s& \xmark & \xmark & 6.7\,s  &   18.9\,s & \xmark\\
3  & Inverted Pendulum & 4 & 10 &   9.7\,s &   0.6\,s& \xmark &112.6\,s& 12.0\,s  &  18.1\,s &  414.0\,s\\
4  & Magnetic Pointer  & 2  & 10 & \xmark & 44.1\,s& \xmark & \xmark& 4.6\,s  &  39.0\,s &  \xmark \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 10 & 14.0\,s & 0.6\,s& 2.0\,s & 1.5\,s &1.2\,s  &   1.5\,s &   4.3\,s\\
6  & Suspension        & 4 & 10 & 73.7\,s& 48.3\,s & \xmark & 110.2\,s &24.0\,s  &   46.1\,s &  557.9\,s\\
7  & Tape Driver       & 3 & 10 & 10.1\,s& 12.4\,s& 140.3\,s &10.5\,s &2.6\,s  &   10.5\,s &   104.6\,s\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results. All times are median (except for Min and Max). \xmark~indicates failure to find a controller (in the median case it means that more than half of the times we failed to find a controller), either due to timeout or insufficient numeric precision (bitvectors are limited to 64bits). -AR1 is standard CEGIS, -AR2 uses Solution Generalization but no refinement. -OR means Abstraction Refinement but no optimization refinement.}
\end{table}

@@

More interesting cases appear when we increase the order, showing initially that
benchmarks that timed-out with AR1 (approx 20mins) are now running in under 2
minutes using AR2. This is the result of removing several multiplications which are
known to be difficult for SAT solvers.  Furthermore, when we introduce Optimization
Refinement there is a further improvement up to an order of magnitude.  This second
improvement is due to the elimination of counterexamples that would have been
created without the refinement. Since further synthesis require increasingly
complex models which are linear in the number of multiplications, eliminating 
these results in large improvements.  In cases when the initial candidate is
suitable, we gain nothing with the refinement and therefore find similar times.
Improvements using this approach will therefore be observed in higher order
systems or when the safety specification is tight and the original loop would
require multiple counterexamples.

@@

>>>>

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Errors generated by numerical representations} 
\label{sec:numeric_rep2}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

In Figure \ref{fig:observersystem2} we introduced several sources of noise,
which are due to quantization effects.  These effects are caused by the use
of imprecise numerical representations over the reals in the digital domain
of the controller which is an important consideration when trying to synthesise
sound controllers.
The models we consider have two sources of error that are due to numerical  
representation which we commonly refer to as FWL effects.
The first is the numerical error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$, $C_d$, $D_d$, $x_k$ and $y_k$.
The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
representation of numbers, and briefly describe the errors
introduced.  A~formal discussion is in
Appendix~\ref{sec:numeric_rep}.

Let $\mathbb{R}_{\langle M,F \rangle}$ be a fixed point domain with $M$ bits
representing the binary word length of the representation and $F$ bits
representing the prescaling $2^{-F}$ (\emph{i.e} the number of bits after
the decimal place).  We note that a negative $F$ denotes an integer
multiplication that adds $-F$ trailing zeros in the number represented by
$M$.

@@

@@

Since we are interested in synthesising controllers with observer
designs, in this section we will look at checking stability, safety
and the response time in the presence of an observer. \footnote{A full description of
the theory related to the observer design is provided in appendix \ref{sec:control_theory}.}

@@

The ability of a system to remain within a bounded set of states is called stability.
In this paper, we are interested in
%% There are different types of stability, such as Bounded Input/Bounded Output (BIBO)
%% in which the requirement is that given a bounded input the output remains within
%% certain limits, or
asymptothic stability, where the output is expected to
converge to a certain value. %% In our model, we are concerned with the latter
%% type of stability is directly related to
%% convergence, which means it directly relates to the eigenvalues of the dynamics.
As we want to use Jury's criterion~\cite{fadali}\footnote{More details on Jury's criterion can be found in~\ref{sec:closed_stability}}, which is expressed 
by an easy algebraic formula, we must ... \note{summarise what we do in the rest of this section.}

@@

The benchmark suite employed to evaluate the proposed technique is 
composed by 7 SISO models extracted from the literature~\cite{acrobot,cstr,CHEN1979389,KOKOTOVIC198023,gajic2008optimal,Franklin15,maglev,converters}.
In order to obtain statistically relevant results, we have taken these models
and modified each possible combination of their parameters by a 10\% tolerance,
thus creating seven hundred benchmarks from this base.
In the verification module, we have used the tool Axelerator~\cite{} .

@@

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{Number of} & \multicolumn{3}{|c|}{CEGIS} & \multicolumn{3}{|c|}{CEGORS} \\
   &                                 & & \multicolumn{1}{|c|}{benchmarks}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max}\\\hline
1  & DC Motor          & 2 & 100 & 9.5\,s & 1.9\,s& \xmark& 1.2\,s &   3.1\,s &   14.1\,s\\
2  & Helicopter        & 3  & 100 & \xmark & \xmark & \xmark & 6.7\,s  &   18.9\,s & \xmark\\
3  & Inverted Pendulum & 4 & 100 &   9.7\,s & \xmark &\xmark & 12.0\,s  &  18.1\,s &  414.0\,s\\
4  & Magnetic Pointer  & 2  & 100 & \xmark & \xmark & \xmark& 4.6\,s  &  39.0\,s &  \xmark \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 14.0\,s & 2.0\,s & \xmark&1.2\,s  &   1.5\,s &   4.3\,s\\
6  & Suspension        & 4 & 100 & 73.7\,s& \xmark & \xmark &24.0\,s  &   46.1\,s &  557.9\,s\\
7  & Tape Driver       & 3 & 100 & 10.1\,s& 140.3\,s & \xmark &2.6\,s  &   10.5\,s &   104.6\,s\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results. Minimum, Maximum and Median times for standard CEGIS vs CEGORS on multiple benchmarks. \xmark~indicates failure to find a controller, either due to timeout or insufficient numeric precision (bitvectors are limited to 64bits).}
\end{table}

@@

More interesting cases appear when we increase the order, showing initially that
benchmarks that timed-out using standard CEGIS are now running in under 2
minutes using CEGORS. This is the result of removing several multiplications which are
known to be difficult for SAT solvers.  Furthermore, when we introduce Optimization
Refinement there is a further improvement up to an order of magnitude.  This second
improvement is due to the elimination of counterexamples that would have been
created without the refinement. Since further synthesis require increasingly
complex models which are linear in the number of multiplications, eliminating 
these results in large improvements.  In cases when the initial candidate is
suitable, we gain nothing with the refinement and therefore find similar times.
Improvements using this approach will therefore be observed in higher order
systems or when the safety specification is tight and the original loop would
require multiple counterexamples.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Modeling quantization as noise} \label{sec:quantization-noise}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
In Figure \ref{fig:observersystem2} we introduced
several sources of noise, which are due to quantization effects. We now describe
their nature and create a sound model for them to be used during our synthesis
and verification process.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Round-off errors}
\label{sec:roundoff-noise}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
